Relative Stable Maps in Genus One via Central Alignments

### 2019-01-01 GenusOneRelativeMaps.tex
Started reorganising and rewriting Section 4

\def\Yagraph{\tikz[baseline=-3pt,scale=.8]{
\draw (2,0) -- (0,1) (2,0) -- (0,.5) (2,0) -- (0,-1);
\draw (2,0) circle(2pt)[fill=black];
\draw (0,1) circle(2pt)[fill=white];
\draw (0,.5) circle(2pt)[fill=black];
\draw (0,-.5) node{$\ldots$};
\draw (0,-1) circle(2pt)[fill=black];
\draw (-1,1.5) -- (1,1.5) -- (1,-1.5) -- (-1,-1.5) -- (-1,1.5);
\draw (1,1.5) -- (3,2) -- (3,-1) -- (1,-1.5);
}}

\def\Ybgraph{\tikz[baseline=-3pt,scale=.8]{
\draw (2,0) to[out=120,in=0] (0,1) (2,0) -- (0,1) (2,0) -- (0,.5) (2,0) -- (0,-1);
\draw (2,0) circle(2pt)[fill=black];
\draw (0,1) circle(2pt)[fill=black];
\draw (0,.5) circle(2pt)[fill=black];
\draw (0,-.5) node{$\ldots$};
\draw (0,-1) circle(2pt)[fill=black];
\draw (-1,1.5) -- (1,1.5) -- (1,-1.5) -- (-1,-1.5) -- (-1,1.5);
\draw (1,1.5) -- (3,2) -- (3,-1) -- (1,-1.5);
}}

\def\Ycgraph{\tikz[baseline=-3pt,scale=.8]{
\draw (2,0) -- (0,1) (2,0) -- (0,.5) (2,0) -- (0,-1);
\draw (2,0) circle(2pt)[fill=white];
\draw (0,1) circle(2pt)[fill=black];
\draw (0,.5) circle(2pt)[fill=black];
\draw (0,-.5) node{$\ldots$};
\draw (0,-1) circle(2pt)[fill=black];
\draw (-1,1.5) -- (1,1.5) -- (1,-1.5) -- (-1,-1.5) -- (-1,1.5);
\draw (1,1.5) -- (3,2) -- (3,-1) -- (1,-1.5);
}}

\newcommand{\mathsq}[1]{#1}
\newcommand{\pM}{\pazocal{M}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\MG}[4]{\overline{\mathcal{M}}^{\rm G}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\VZ}{\pazocal{V\!Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\VZrel}[4]{\mathcal{V\!Z}^{\rm{rel}}_{#1,#2}(#3,#4)}
\newcommand{\stab}{\rm{stab}}
\newcommand{\stC}{C'}
\newcommand{\stf}{f'}
\newcommand{\stpi}{\pi'}
\newcommand{\sts}{s'}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\fgt}{\operatorname{fgt}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\Acal}{\mathcal{A}}
\newcommand{\Mcal}{\mathcal{M}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\gp}{\operatorname{gp}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

In the previous section we used tropical geometry to identify the logarithmic strata of $\VZ_{1,\alpha}(\PP^N|H,d)$. We now focus our attention on the logarithmic divisors, providing a recursive description for these loci: i.e. a description using a fibre product of smaller moduli spaces. This is the technically most difficult part of the paper. As a result, we obtain a Gathmann recursion formula in genus one.

Note that $f$ has contact order with $H$ exactly equal to $\alpha_k$ at the marking $p_k$, or else the irreducible component of $C$ on which $p_k$ lies is mapped entirely inside $H$. By pulling back along $f$ the equation defining $H$, and taking its $\alpha_k$-th derivative at $p_k$ (which makes sense because all the lower order derivatives do vanish by assumption) we single out the latter locus. By staring at the exact sequence of jet bundles
\[0\to p_k^*\Omega_{C}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)\to p_k^*\mathcal J^{\alpha_k}(f^*\OO_{\PP^N}(H))\to p_k^*\mathcal J^{\alpha_k-1}(f^*\OO_{\PP^N}(H))\to 0\]
we realise that there is on $\VZ_{1,\alpha}(\PP^N|H,d)$ a line bundle (the leftmost term in the exact sequence above) with a natural section cutting the locus where the curve breaks, and the piece containing $p_k$ is mapped inside $H$.
\begin{lem}
 \[(\alpha_k\psi_k+\ev_k^*H)[\VZ_{1,\alpha}(\PP^N|H,d)]=[D_{1,\alpha;k}(\PP^N|H,d)],\]
\end{lem}
In what follows we shall give an explicit (recursive) description of the right-hand side, in terms of tautological classes on spaces of maps with lower numerical invariants.
\begin{remark}
 The divisors we are after are a union of logarithmic strata, because the locus where the log structure is trivial coincides precisely with the nice locus.
\end{remark}
By Corollary \ref{cor:log_smooth}, the logarithmic strata of codimension one correspond to rays of the Artin fan.

\subsection{Combinatorial description}
As long as the minimal subcurve of genus one has positive degree, the picture is classical: the dual graph of the curve is bipartite - vertices mapped inside $H$ are distinguished from the others. In fact there is only one such vertex, because, if there were more, the relative position of their image in $\mathbb R_{\geq 0}$ under the tropical map would be unconstrained, therefore spanning a locus larger than a ray in the tropical moduli space. \marginpar{say something about rigid tropical curves as in ACGS and KLR} We follow in R. Vakil's footsteps in setting the following notation:
\begin{figure}[h]
    \centering
    \begin{minipage}{0.3\textwidth}
        \centering
        \Yagraph
        \caption{$\mathcal Y_a$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ybgraph
        \caption{$\mathcal Y_b$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ycgraph
        \caption{$\mathcal Y_c^+$}
    \end{minipage}
\end{figure}

Notice that continuity of the tropical map determines the mutual relationship among the edge lengths.

On the other hand, when there is a contracted elliptic subcurve - and it will be contracted into the hyperplane, because otherwise it wouldn't be generic, by density of the nice locus in $\VZ_{1,\alpha^\prime}(\PP^N|H,d^\prime)$ - the picture becomes more complicated due to the alignment. The combs may break (multiple times) and they may point in opposite directions. We label these loci $\mathcal{Y}_c^0$.
\begin{figure}
 \tikz{
 \draw (0,0) circle(2pt)[fill=black] (0,2) circle(2pt)[fill=black] (0,4) circle(2pt)[fill=black] (0,6) circle(2pt)[fill=black] (0,-2) circle(2pt)[fill=black] (1.1,1.78) circle(2pt)[fill=black] (3.42,1.42) circle(2pt)[fill=black] (.58,-1.42) circle(2pt)[fill=black];
 \draw (2,0) -- node[above]{4} (0,0) (2,0) --node[above]{2} (1.1,1.78) --node[above]{2} (0,2) (1.1,1.78) --node[above]{1} (0,4) (2,0) --node[above]{1} (3.42,1.42) (2,0) --node[above]{1} (.58,-1.42) (3.42,1.42) --node[above]{5} (0,6) (.58,-1.42) --node[above]{3} (0,-2);
 \draw[red] (2,0) circle(2);
 \draw (2,0) circle(2pt)[fill=white];
 \draw[->] (0,-3) circle(2pt)[fill=black] edge (4,-3);
 }
\end{figure}
\begin{dfn}
 A degree-aligned genus one tropical map $\phi$ to $\mathbb R_{\geq 0}$ is ... with a circle of radius $\delta$ around the core satisfying:
 \begin{enumerate}
  \item the circle of radius $\delta$ passes through at least one vertex of positive degree, and no vertex of positive degree is contained in its strict interior;
  \item every circle of radius $<\delta$ around the core passes through (either none or) at least two vertices;\footnote{The subdivisions we perform before picking a ray correspond to aligning the log structure, i.e. imposing $\lambda(v)=\lambda(v^\prime)$ for some $v,v^\prime$.}
  \item every edge heading out of the circle of radius $\delta$ goes directly to a vertex in $\phi^{-1}(0)$.
 \end{enumerate}
\end{dfn}

\subsection{Explicit realisation}
The description of the first three types is relatively easy:
\begin{enumerate}[label=(\alph*)]
 \item \[\mathcal Y_a=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^N|H}{d_1}\times\prod_{i=2}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_b=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^N|H}{d_1}\times\prod_{i=3}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_c^+= \VZc{1}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^{\sim} \times_{H^r}\prod_{i=1}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\]
\end{enumerate}
where $\M{0}{\mu}{H}{d}^\sim$ denotes the space of genus zero punctured maps, while $\VZc{0}{\mu}{H}{d}$ denotes the desingularisation of the main component of the space of punctured maps discussed in Section ???. The numerical data run over all splitting $A$ of the tangency conditions (resp. $B$ of the degree) in $r+1$ parts, such that:
\begin{enumerate}
 \item $d_0+\sum_{i=1}^r m^{(i)}=\sum\alpha^{(0)}$;
 \item $d_i=m^{(i)}+\sum\alpha^{(i)}$;
 \item $d_0>0$ in case $c$.
\end{enumerate}

>>>>

% FIGURES FOR USE LATER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\Yagraph{\tikz[baseline=-3pt,scale=.8]{
\draw (2,0) -- (0,1) (2,0) -- (0,.5) (2,0) -- (0,-1);
\draw (2,0) circle(2pt)[fill=black];
\draw (0,1) circle(2pt)[fill=white];
\draw (0,.5) circle(2pt)[fill=black];
\draw (0,-.25) node{$\vdots$};
\draw (0,-1) circle(2pt)[fill=black];
\draw [red,fill=red] (0,-2) circle[radius=2pt];
\draw [->,red,thick] (0,-2) -- (2,-2);
\draw (2,-2) node[right,red]{\tiny{$\Sigma(\PP^N)$}};
\draw [->,red] (1,-0.9) -- (1,-1.6);
}}

\def\Ybgraph{\tikz[baseline=-3pt,scale=.8]{
\draw (2,0) to[out=120,in=0] (0,1) (2,0) -- (0,1) (2,0) -- (0,.5) (2,0) -- (0,-1);
\draw (2,0) circle(2pt)[fill=black];
\draw (0,1) circle(2pt)[fill=black];
\draw (0,.5) circle(2pt)[fill=black];
\draw (0,-.25) node{$\vdots$};
\draw (0,-1) circle(2pt)[fill=black];
\draw [red,fill=red] (0,-2) circle[radius=2pt];
\draw [->,red,thick] (0,-2) -- (2,-2);
\draw (2,-2) node[right,red]{\tiny{$\Sigma(\PP^N)$}};
\draw [->,red] (1,-0.9) -- (1,-1.6);
}}

\def\Ycgraph{\tikz[baseline=-3pt,scale=.8]{
\draw (2,0) -- (0,1) (2,0) -- (0,.5) (2,0) -- (0,-1);
\draw (2,0) circle(2pt)[fill=white];
\draw (0,1) circle(2pt)[fill=black];
\draw (0,.5) circle(2pt)[fill=black];
\draw (0,-.25) node{$\vdots$};
\draw (0,-1) circle(2pt)[fill=black];
\draw [red,fill=red] (0,-2) circle[radius=2pt];
\draw [->,red,thick] (0,-2) -- (2,-2);
\draw (2,-2) node[right,red]{\tiny{$\Sigma(\PP^N)$}};
\draw [->,red] (1,-0.9) -- (1,-1.6);
}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\mathsq}[1]{#1}
\newcommand{\pM}{\pazocal{M}}
\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\MG}[4]{\overline{\mathcal{M}}^{\rm G}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\VZ}{\pazocal{V\!Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\VZrel}[4]{\mathcal{V\!Z}^{\rm{rel}}_{#1,#2}(#3,#4)}
\newcommand{\stab}{\rm{stab}}
\newcommand{\stC}{C'}
\newcommand{\stf}{f'}
\newcommand{\stpi}{\pi'}
\newcommand{\sts}{s'}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\fgt}{\operatorname{fgt}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\Acal}{\mathcal{A}}
\newcommand{\Mcal}{\mathcal{M}}
\newcommand{\Dcal}{\mathcal{D}}
\newcommand{\Ycal}{\mathcal{Y}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\gp}{\operatorname{gp}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\noindent In the previous section we used tropical geometry to identify the logarithmic strata of $\VZ=\VZ_{1,\alpha}(\PP^N|H,d)$. We now focus our attention on the logarithmic divisors, providing a recursive description for these loci: that is, a description in terms of fibre products of smaller moduli spaces. This is the technically most difficult part of the paper. The recursive description bears some similarities with the recursive desription of the boundary of the absolute space $\VZ_{1,n}(\PP^N,d)$ given in \cite{VZ}; in keeping with the theme of this paper, however, we will see that the logarithmic structures provide a convenient framework for dealing with the combinatorics.

The end result is that we obtain an explicit formula for tautological integrals over these boundary loci. This produces a Gathmann recursion formula for relative invariants in genus one.

\subsection{Recursion setup} Consider an element of $\VZ$ and choose a marked point $x_k\in C$. We note that by the existence of a log morphism, either $f$ has contact order $\alpha_k$ to $H$ at the marking $x_k$, or else the irreducible component of $C$ containing $x_k$ is mapped entirely inside $H$ (we say ``$x_k$ belongs to an internal component'' in this case).

By pulling back the equation defining $H$ along $f$ and then taking its $\alpha_k$-th derivative at $x_k$ (which makes sense since all the lower-order derivatives vanish by assumption) we obtain a section of a jet bundle on $\VZ$ whose vanishing locus coincides (set-theoretically) with the locus where $x_k$ belongs to an internal component. As in \cite[Construction 2.1]{Ga} there is an exact sequence of jet bundles
\begin{equation*} 0\to x_k^*\omega_{\mathcal{C}/\VZ}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)\to x_k^*\mathcal J^{\alpha_k}(f^*\OO_{\PP^N}(H))\to x_k^*\mathcal J^{\alpha_k-1}(f^*\OO_{\PP^N}(H))\to 0 \end{equation*}
and thus we obtain a section of the line bundle $x_k^*\omega_{\mathcal{C}/\VZ}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)$ with vanishing locus equal to the locus where $x_k$ belongs to an internal component. Writing this as
\begin{equation*} \Dcal^k = \Dcal_{1,\alpha}^k(\PP^N|H,d) \subseteq \VZ_{1,\alpha}(\PP^N|H,d) = \VZ \end{equation*}
we obtain:
\begin{lem} $(\alpha_k\psi_k+\ev_k^*H)\cap[\VZ]=[\Dcal^k].\footnote{(Navid) Should we work out the vanishing orders of the section at this point?}$\end{lem}
\noindent The remainder of this section will be devoted to providing an explicit (recursive) description of $[\Dcal^k]$ in terms of tautological classes on spaces of maps with smaller numerical invariants. As we will see, this provides an algorithm for computing reduced relative invariants, as well as a quantum Lefschetz algorithm for reduced (absolute) invariants in genus one.

\subsection{Recursive description of the divisors: types $A,B$ and $C^+$}
We begin by noting that every irreducible component of $\Dcal^k$ is a codimension--$1$ logarithmic stratum. This is obvious: the locus where the log structure is trivial coincides precisely with the nice locus, where the source curve is smooth and not mapped inside $H$. By definition $\Dcal^k$ is disjoint from this locus, and since it has codmension $1$ it must be equal to a union of logarithmic strata.

As discussed at length in the previous section, the divisorial logarithmic strata of $\VZ$ correspond to the rays of the polyhedral subdivision of the tropical moduli space which produces the Artin fan $\Acal$. The logarithmic strata contained in $\Dcal^k$ correspond precisely to the rays of the tropical moduli space where the vertex of the tropicalisation containing $x_k$ is mapped into the interior of $\RR_{\geq 0}$ by the tropical map.

We proceed as in the previous section, fixing an element of $\Mcal$ and considering the resulting subdivision of the tropical moduli space. As long as the minimal subcurve of genus one is not contracted, the picture is classical (i.e. the subdivision is trivial, and $\VZ=\Mcal$ locally). In this case there may only be one vertex of the tropical curve mapped into the interior of $\RR_{\geq 0}$; if there were more, the positions of their images in $\RR_{\geq 0}$ would be unconstrained and thus the corresponding maximal cone in the tropical moduli space would be at least $2$--dimensional, producing a locus in $\VZ$ of larger codimension.\footnote{(Luca) Say something about rigid tropical curves as in ACGS and KLR} We adopt the terminology of R. Vakil, distinguishing three cases as below:

\begin{figure}[h]
    \centering
    \begin{minipage}{0.3\textwidth}
        \centering
        \Yagraph
        \caption{$\mathcal Y_A$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ybgraph
        \caption{$\mathcal Y_B$}
    \end{minipage}\hfill
    \begin{minipage}{0.3\textwidth}
        \centering
        \Ycgraph
        \caption{$\Ycal_C^+$}
    \end{minipage}
\end{figure}
\noindent Notice that the continuity of the tropical map identifies all of the edge lengths up to scalars given by the expansion factors (i.e. tangency orders) at the edges; thus the tropical moduli space is isomorphic to $\RR_{\geq 0}$ and so we have a divisor (with no subdivision required).

The description of the first three types is relatively easy:
\begin{enumerate}[label=(\alph*)]
 \item \[\mathcal Y_a=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^N|H}{d_1}\times\prod_{i=2}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_b=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^N|H}{d_1}\times\prod_{i=3}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_c^+= \VZc{1}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^{\sim} \times_{H^r}\prod_{i=1}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\]
\end{enumerate}
where $\M{0}{\mu}{H}{d}^\sim$ denotes the space of genus zero punctured maps, while $\VZc{0}{\mu}{H}{d}$ denotes the desingularisation of the main component of the space of punctured maps discussed in Section ???. The numerical data run over all splitting $A$ of the tangency conditions (resp. $B$ of the degree) in $r+1$ parts, such that:
\begin{enumerate}
 \item $d_0+\sum_{i=1}^r m^{(i)}=\sum\alpha^{(0)}$;
 \item $d_i=m^{(i)}+\sum\alpha^{(i)}$;
 \item $d_0>0$ in case $c$.
\end{enumerate}
